---
title: "Initial_Explore"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Initial exploration of EWE data for Eco-DAS Extreme Events project

AGH, 13Nov19

Load data and toolboxes. Create dataset that includes single unique accessions

```{r}
pacman::p_load(tidyverse,ggplot2)

data <- read.csv('C:/Users/ahoun/Dropbox/EWE/EWE/Data/CleanedExtremeEventsData_8Nov2019.csv')
data_uni <- distinct(data,UniqueAccession,.keep_all=TRUE)
data_uni <- data_uni[-50,]

```

## Are these studies linking to climate?

First count responses as 'y' and 'n'

```{r}
climate <- data_uni %>% group_by(Climate_relationship) %>% tally()

```

## Plot as a pie chart

```{r pressure, echo=FALSE}
lbls <- c("No Response","No","Yes")
pie(climate$n,labels=lbls,main="Climate relationship?")
```

## What parameters are these studies measuring?

Biological? Physical? Chemical?
Used whether Raw Data was available for each category: would need to update for actual analysis!

```{r}
bio <- data_uni %>% group_by(BiologicalResponse_RawData_PublicallyAvailable) %>% tally()
chem <- data_uni %>% group_by(ChemicalResponse_RawData_PublicallyAvailable) %>% tally()
phys <- data_uni %>% group_by(PhysicalResponse_RawData_PublicallyAvailable) %>% tally()
ana <- matrix(c(38,11,25,11,38,24),ncol=3,byrow=TRUE)
colnames(ana) <- c("Bio","Chem","Phys")
rownames(ana) <- c("Yes","No")

```

## Plot as a bar chart

```{r pressure, echo=FALSE}
barplot(ana,legend=TRUE,beside=TRUE)
```
